---
title: "Introduction to selectiveMLE"
output:
  rmarkdown::github_document
vignette: >
  %\VignetteIndexEntry{Introduction to selectiveMLE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`selectiveMLE` provides tools for conditional maximum likelihood inference
when a model has been chosen using the lasso or via thresholding of normal
means.

This vignette demonstrates basic usage of the two exported functions
`lassoMLE()` and `truncNormMLE()`.

## Lasso example

The code below follows the small simulation from the README.  We generate a
matrix of predictors and a response vector, fit a lasso model and then obtain
the conditional MLE for the selected coefficients.

```{r lasso, message=FALSE}
library(selectiveMLE)
set.seed(123)
X <- matrix(rnorm(100 * 10), 100, 10)
beta <- c(2, -1.5, rep(0, 8))
y <- as.numeric(X %*% beta + rnorm(100))

fit <- lassoMLE(y, X, lambda = "lambda.min", optimSteps = 200,
                sampSteps = 400, verbose = FALSE)
fit$conditionalBeta
fit$wald_CI
sol <- data.frame(iter = seq_len(nrow(fit$solution_path)),
                  fit$solution_path)
sol_long <- tidyr::pivot_longer(sol, -iter,
                                names_to = "variable",
                                values_to = "estimate")
ggplot(sol_long, aes(iter, estimate, colour = variable)) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw()

knitr::kable(fit$wald_CI, col.names = c("lower", "upper"), digits = 2)
```

## Selected normal means

The package also handles inference for truncated normal observations via
`truncNormMLE()`.  Here we generate a small example and compute the MLE
for coordinates that cross a threshold of one.

```{r mvtnorm, message=FALSE}
sigma <- diag(5)
y <- rnorm(5, mean = 0, sd = 1.5)
res <- truncNormMLE(y, sigma, threshold = 1,
                    maxiter = 300, verbose = FALSE)
res$mle
res$CI
barplot(rbind(observed = y, mle = res$mle), beside = TRUE,
        legend.text = TRUE, ylab = "value")
```

More elaborate demonstrations can be found in the scripts
`lassoMLE example script.R` and `mvtnorm example script.R` located in the
repository root.
